### 2019-04-16 papers/SymmetricLenses/rebuttal_icfp19.txt

Thank you for your feedback. We respond the major questions and concerns, then
provide an (optional) point-by-point response to individual concerns. We are
happy to include any of the information presented in this rebuttal in the final
version of the paper.

    - Why is writing two regexes easier for the user than writing a lens? Do
      input formats often come equipped with regexes already? Can the
      synthesizer use such pre-existing regexes out of the box, or do they have
      to be "massaged" by the user?

While we require users to write regular expressions, programming regular
expressions is easier than programming lenses:

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while a lens deals with two
  at the same time.

* In the (common) case where the two formats have Kleene stars in different
  places, manual lens programming requires mentally transforming both to a
  common "aligned" form.  Our synthesis algorithm does this alignment
  automatically.  This alignment can be quite challenging for programmers.

As an explicit example, consider the following function:

        let last_head = NAME . "," . (WSP . NAME)*
        let last_tail = (NAME . WSP)* NAME

        let name_conversion : last_head <=> last_tail =
            lens_swap
                (NAME . del ",")
                (lens_swap WSP NAME)*

While the name_conversion expression may have more AST nodes than the two
regular expressions, we find it much harder to write.

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not require these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, the types provide resilience in the face of future
lens modifications, as the types ensure that the lens maps exactly
between strings of the provided formats. Invalid inputs are not accepted
nor are the outputs ever invalid.  Consequently, we are not requiring
additional work. Instead, we are saving programmers from having to 
writing the lens by leveraging work they should already be performing.

In our experiments, we manually programmed the regular expressions. While
writing these regular expressions, we did not massage them to be amenable to
synthesis. In general, the synthesizer is able to handle regular expressions
expressed in multiple different ways, due to the traversal of regular expression
equivalences. However, if the regular expressions require a transformation that
was not a star-semiring transformation, such a massaging would be required.

    - What is the relationship between our information-theoretic ranking, and
      prior work like Unsupervised Learning by Program Synthesis

(afm: this can potentially go after the fold, as it was only mentioned in
passing, not as a critical question of the reviewer)

Unsupervised Learning by Program Synthesis (ULPS) expands upon previous research
in unsupervised machine learning. Unsupervised learning attempts to "induce good
latent representations of a data set" [Ellis et al. 2015]. Given a set of data
points, or "outputs," ULPS attempts to find good representations of these
outputs as a combination of finding a single function, and a sequence of
"inputs" that correspond to the outputs. The latent representation is a
combination of the corresponding input, and the transformation from inputs to
outputs.

Our work is similar to this work in that we both use information-theoretic
principles to find the desired programs, but the use of information theory in
each is different. In ULPS, minimizing the information theoretic cost metric is
the only goal. In this metric, the desire is to encode the output data in as few
bits as possible, so the example data is tightly coupled with the problem. In
our work, we use information theory to formalize the notion of "more bijective,"
providing a firm theoretical foundation for previous ad-hoc ranking functions.
In our case, the examples serve merely as a specification, but the inclusion or
exclusion of such examples does not impact "how bijective" certain functions
are. We do not consider this a weakness, but an inherent aspect to our
formalization. Furthermore, we would not consider it a good idea to assume that
we should try to minimize the data recovery costs on the examples more than on
other pieces of data -- non-bijective input/output pairs are often useful.

    - The authors do not consider lens compositions in their synthesis. Are
      there any justifications not to consider them? In existing results that
      synthesize bijective (quotient) lenses, one of the reasons not to consider
      lens composition in their synthesis is that ruling out them does not
      change the expressive power? Can you say a similar argument for this case?

We do not synthesize lens compositions for the same reasons as previous work --
such compositions are quite difficult to generate. You are correct that we do
not prove that such a restriction does not limit the expressive power. However,
we do not consider the symmetric DNF lens language a core contribution
(symmetric DNF lenses are not even defined outside of the appendix) and do not
believe such a proof would add much to the paper. However, we do provide
experimental evidence that the composition operator is not needed in practice;
we are able to synthesize all our benchmarks.

    - There is no explanation of the selection of the benchmarks. What are their
      characteristics and why those exercise interesting aspects of the
      algorithm.

We describe the selection of the benchmark suite in section 6.1. 29 of the
benchmarks are taken from Augeas, a bidirectional tool for editing Linux
configuration files. These examples were quite large, and involved some very
complex transformations, though these benchmarks were typically bijective. These
demonstrated how our tool performed at large, complex transformations. The 8
benchmarks from Flash Fill provided a range of transformations, none of which
were not bijective. These demonstrated how our tool performed on tasks that lost
large amounts of data, and sometimes required extended transformations through
Expand. The remaining 11 benchmarks tested how our tool performed at the type of
transformations we are familiar with, from transformations that would be useful
in day-to-day software engineering or classic transformations from the
bidirectional programming literature.

    - There is no attempt to explain what are the sources of worse performance
      for SS.

The primary impact on speed is the use of GreedySynth, instead of the bijective
RigidSynth. For GreedySynth, lenses are found between regular expressions that,
ultimately, do not show up in the final lens. RigidSynth does not incur this
performance penalty -- it only synthesizes lenses between regular expressions
for which there exists a bijective lens between them.

=================================================================================
Reviewer A
=================================================================================

    - Moreover, on top of the regexes, the user needs to provide examples and
      relevance annotations, and the only way to do so seems to be through an
      iterative process (by inspecting the synthesized lens and adding more
      examples / annotations when needed). Such iterative approach would be
      challenging for a user who is not sufficiently familiar with lenses to
      write one themselves.

We described the iterative process assuming a lazy user who didn't want to
provide more specifications than were necessary. However, such an iterative
process is not required, and if the user is confident that certain information
is "required" or may safely "skipped," they can safely add this information --
it will not slow down synthesis. Similarly, if the user already has a benchmark
suite, they may safely include such examples.

    - To add to the previous point: the evaluation could do a better job
      quantifying these limitations. For example, it would be good to see for
      each benchmark the size of annotations required, as well as how many
      "existing lenses" were used in compositional synthesis and where they came
      from (some standard library of lenses? or did the authors manually
      decompose each problem into subproblems? if so, then how many subproblems
      were there?)

There were a total of 12 uses of "require" in the benchmark suite. There were a
total of 4 uses of "skip" in the benchmark suite. In total 73 existing lenses
were used in compositional synthesis (about 1.5 per benchmark on average).
Typically, we manually decomposed these lenses, though some used the to_upper
function. We are happy to include a figure detailing the subtasks specified
during compositional synthesis in the final version of the paper (similar to the
one on p24 of Synthesizing Bijective Lenses).

    - Finally, the evaluation could also do a better job confirming that there's
      indeed a benefit to synthesizing a lens, rather than using a tool like
      FlashFill to synthesize the two functions independently. The related work
      section mentions that FlashFill cannot handle nested loops. However,
      FlashFill has evolved a lot since its original publication in 2011, so it
      would be interesting to see an empirical comparison.

We included an extended comparison to FlashFill in previous work, and did not
include such a comparison for lack of space. We are happy to make comparison to
FlashFill cited more explicitly.

    - The introduction mentions that stochastic regular expressions guide the
      search process. I would like to see it better explained in the main text
      of the paper: do they actually guide the search or merely rank the
      results?

Due to the Greedy nature of GreedySynth, the stochastic regular expressions
guide the search (but the stochastic annotations only guide the search insofar
as the lens costs guide the search).

    - L 284: please explain why this rule is needed for the lens in Fig. 3 to be
      well-typed.

Without the type equivalence rule, the lens would be well-typed between
emp_salaries and header . ("" | ("\n" . emp_ins) . ("\n" . emp_ins)∗). To be
well-typed between emp_salaries and emp_insurance, the right-hand type needs to
be transformed into an equivalent form.

    - Section 5.6: expansion inference seems to be a crucial part of the
      algorithm but is described very briefly. I would like to see more detail.

Expansion inference is the same as is performed in bijective lens inference, and
we did not include it for space. We will gladly make such a citation more
explicit.

    - Can you explain why you need the lower bound [for intervals]?

While we do not need it in the algorithm, we include it to make formal claims
about bounding the cost. We describe the need for intervals in 4.1.

### 2019-04-17 report/final-report.tex
edits

\documentclass[12pt]{article}

@@

@@

  \large
  %\begin{tabular}{rl}
  Kathleen Fisher (Tufts University) \\
  Benjamin Pierce  (University of Pennsylvania) \\
  David Walker (Princeton University) \\
  Steve Zdancewic (University of Pennsylvania)
 %\end{tabular}
  \vspace{1in}

@@

\tableofcontents
\newpage

\listoffigures
\newpage

@@

\section{Introduction}

Building information-processing systems and maintaining them over a long
period of time is a tedious, labor-intensive process.  One key challenge is
that such systems must often interact with a large number of
\emph{ad hoc data sources}---partially structured data sources represented
in non-standard formats.  Ad hoc data sources include various
different kinds of system log files as well as scientific data sources
generated by experiments.  These ad hoc data sources are often
produced by other automated systems, and each requires custom tools.
Over time, the data sources tend to evolve---fields are added,
removed, or co-opted, variants are added, etc.  Today, to manage these
changes, engineers must manually re-code parsers and/or insert
adaptors while maintaining the desired semantics.  Such manual work is
not only time-consuming, but exceedingly error-prone.  Moreover,
errors in environment-facing interfaces can not only lead to
corruption of important data, but also to significant security
vulnerabilities.  In order to design and implement survivable,
long-lived, complex software systems that are robust to changes 
in their operational environment---the vision of
DARPA's BRASS program---it is necessary to develop new, easier-to-use
and more robust programming systems for managing these ad hoc data
sources as they evolve.

@@

The core result of our research is a new algorithm for
the synthesis of \emph{bijective string lenses}.  Bijective string lenses (\emph{i.e.,}
bidirectional transformations) define a limited set of
transformations between strings.  The domain and range of such
transformations are determined by regular expressions ({\emph{i.e.,}
  regular expressions serve as the types of these transformations).
 In addition, as their name suggests, these transformations are
 \emph{bijections}.
 In other words, the information content of source A is preserved (though usually
rearranged) when data is translated to target B, and vice versa when B
is translated back to A.  Such transformations can rearrange fields of
a record or insert new kinds of syntactic separators (\emph{e.g.,} replacing a
comma with a vertical bar, or the name of one HTML tag with another)
but they cannot implement more general transformations that elide
irrelevant details, such as the amount of whitespace that separates
two tokens.

@@

\section{Methods, Assumptions and Procedures}

@@

Lens-based languages are present in variety of tools and have found mainstream
industrial use.
Boomerang~\cite{boomerang, Matching10} lenses provide
guarantees on transformations between {\em ad hoc} string document formats.
Augeas~\cite{augeas}, a popular tool that reads Linux system configuration
files, uses the \emph{get} part of a lens to transform configuration
files into a canonical tree representation that users can edit
either manually or 
programmatically.  It uses the lens's \emph{put} to merge the edited
results back into the original string format.  Other lens-based languages and
tools include 
%
GRoundTram~\cite{Hidaka2011GRoundTramAI},
%
BiFluX~\cite{DBLP:conf/ppdp/PachecoZH14}, 
%
BiYacc~\cite{DBLP:conf/staf/ZhuK0SH15},
%
Brul~\cite{DBLP:conf/etaps/ZanLKH16},
%
bidirectional variants of 
relational algebra~\cite{BohannonPierceVaughan},
%
BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
%
spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
graph query languages~\cite{DBLP:conf/icfp/HidakaHIKMN10},
and
XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.

If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
$\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
stochastic regular expressions, so we mark plain REs with overbars.)
We illustrate some of these
combinators by defining lenses on subcomponents of the employee data formats.

@@

% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens is useful for synchronizing a series of items or rows in a
table. For example given a
lens \lstinline{employee_lens} that synchronizes data for a single employee, the
lens
%
\begin{lstlisting}
iterate(id("\n") . employee_lens)  :  ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%
transforms a list of employees in employees in the Management format to a list
of employees in the HR format and vice versa. These combinators
are combined in figure~\ref{fig:example_lens} to construct a complete lens
between the employee formats.

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens that synchronizes management and HR employee files}
  \label{fig:example_lens}
\end{figure}

@@

  \end{tikzpicture}
  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
    \RegexAlt{}, and examples, \Examples{}, are given as input.
    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
    \RegexAlt{}, and \Examples{}.
    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
    that is equivalent to \DNFLens{}.}
  \label{fig:schematic-diagram-synthesis}
\end{figure}

@@

\section{Results and Discussion}

@@

\paragraph*{Benchmark Suite Construction}
We constructed our benchmarks primarily by adapting examples from
Augeas~\cite{augeas} and 
Flash Fill~\cite{gulwani-popl-2014}.
%
Augeas is a configuration editing system for Linux that uses lens
combinators similar to those in Boomerang.  
% , with some small
% differences. However, it transforms 
% strings on the left to structured trees on the right rather than
% transforming strings to strings.
% We adapted these Augeas lenses to our setting by converting the
% right-hand sides to strings that correspond to serialized versions
% of the tree formats
We derived 29 of the benchmark tests by
adapting the first 27 lenses from the Augeas library in alphabetical order,
as well as the lenses 
\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
by the `A' lenses.
Furthermore, the 12 last synthesis problems derived
from Augeas were tested after our synthesizer was was
finalized, demonstrating that the optimizations were not
over-tuned to perform well on the testing data.
%
Flash Fill is a system that allows users to specify common string
transformations by example~\cite{gulwani-popl-2014}.  
We derived three benchmarks from the first few examples in this
paper and one from the running example on
extracting phone numbers.

@@

Figure~\ref{fig:definition-sizes} shows the complexity of our type-based
specifications as well as our example counts.
An average benchmark has a type-based specification that can be represented using
310 AST node, and requires 1.1 input/output examples.
Our benchmarks vary from simple problems, like changing the
representation of dates
(with a specification size of 85, and a generated lens size of 79), to 
complex tasks, like transforming configuration files for server monitoring
software into dictionary form (with a specification size of 670 and
a generated lens size of 651).  On average, the size of the generated lens is
89\% the size of its type specifications.

@@

We analyzed the same 39 lens synthesis tasks as we did for the original
Optician system.
We also experimented using our tool to synthesize quotient lenses
between XML, RDF and and JSON formats using data from the data.gov
database; the data consisted of census statistics, demographic
statistics, wage comparosion data, and crime index data).
Recall that, to use bijective synthesis, we had to modify some of the 
benchmarks (10 of 39) to make them bijective---for instance, by normalizing whitespace.
Because quotient lenses are more expressive, we were able avoid such
modifications. This
experience alone points to the benefits that quotient lens synthesis
brings to the table.

@@

We are interested in comparing three different approaches, which vary
in the amount of synthesis used. 
In the first approach, which we call \QRESize{}, the programmer uses
quotient lens synthesis.  She must write QRE specifications of the source and
target formats and she may give examples.
In the second approach, which we call \canonizeAndSpecSize{} for
Bijective Synthesis, the
programmer uses bijective lens synthesis \`a la Optician.
She must write canonizers by hand, along with 
regular expressions to describe the external
representations of the source and target formats. (The internal
formats can be inferred from the canonizers.) She may also
provide examples to help in the synthesis of the bijective lens.
In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
programmer writes the lens between the source and target formats
entirely by hand, including the descriptions of the source and target
formats.

For each problem in the benchmark suite, we calculate the following measures
as proxies for the level of programmer effort when using each the three
approaches:

%
\begin{itemize}
  \item[\QRESize{}:] 
  The number of AST nodes in the QRE specifications for the source and
  target formats, including examples. 
  \item[\canonizeAndSpecSize{}:] 
  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$
  requiring a  non-trivial canonizer (``canonizers'' as transformations that
  project a string to its equivalence class representative; they are written as
  programs in Boomerang), and (3) the number of AST nodes in the
  examples.  We use (1) to estimate the burden of describing
  the external source and target formats and (2) to estimate the
  burden of writing the requisite canonizers
  by hand.  We count the nodes in the examples because they would be
  fed to the bijective synthesizer.  
  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
  automatically generated from the corresponding QRE $q$, and it is
  possible that a human-written version might be smaller.
  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
  number of AST nodes in the synthesized QRE lens.  We use (1) to
  estimate the burdern of describing the source and target formats
  and (2) to estimate the burdern of writing the appropriate lens by
  hand. These counts are also approximations, as
  $W(q)$ and the synthesized lens may be larger than one written by hand.
\end{itemize}

@@

\subsection{Simple Symmetric Lens Synthesis}
In addition to the results above, we implemented simple symmetric lenses as an extension to
Boomerang~\cite{boomerang}.  
This allows users
to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into manually-written
lenses, and reference previously defined lenses during synthesis. All experiments
were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of 1600 MHz DDR3
running macOS Mojave.

@@

\begin{figure}
  \centering
  \includegraphics{times}
  \vspace{-2ex}
  \caption{Number of
    benchmarks that can be solved by a given algorithm in a 
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. Without compositional
    synthesis it is able to complete 26. Each benchmark specification includes
    source and target (potentially annotated) regular expressions, and between
    one and three sufficient examples.}
  \label{fig:times}
\end{figure}

For each benchmark in the suite and each mode, we ran the system with a timeout
of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:times}
summarizes the results of these tests. We found that our algorithm is able to
synthesize all of the benchmarks in under 30 seconds. Without compositional
synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.

@@

\section{Conclusions}

@@

\bibliographystyle{apalike}
\bibliography{local.bib,bcp.bib}

>>>>

\documentclass[12pt]{article}
%\usepackage[style=ieee]{biblatex}
%\addbibresource{local.bib,bcp.bib}
%\documentclass{IEEEtran}
\usepackage{etoolbox}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}

\providetoggle{blx@skipbiblist}

@@

\newcommand{\FS}{\vspace{10pt}}

@@

  \large
  %\begin{tabular}{rl}
  Kathleen Fisher (Tufts University) \\
  Solomon Maina (University of Pennsylvania) \\
  Anders Miltner (Princeton University) \\
  Benjamin Pierce  (University of Pennsylvania) \\
  David Walker (Princeton University) \\
  Steve Zdancewic (University of Pennsylvania)
 %\end{tabular}
  \vspace{1in}

@@

\addcontentsline{toc}{section}{{Contents}}
\tableofcontents
\newpage

%\cleardoublepage
% \phantomsection
\addcontentsline{toc}{section}{{List of Figures}}
\listoffigures

\newpage

@@

\newpage
\section{Introduction}

Building information-processing systems and maintaining them over a long
period of time is a tedious, labor-intensive process.  One key challenge is
that such systems must often interact with a large number of
\emph{ad hoc data sources}---partially structured data sources represented
in non-standard formats.  Ad hoc data sources include various
different kinds of system log files as well as scientific data sources
generated by experiments.  These ad hoc data sources are often
produced by other automated systems, and each requires custom tools.
Over time, the data sources tend to evolve---fields are added,
removed, or co-opted, variants are added, etc.  Today, to manage these
changes, engineers must manually re-code parsers and/or insert
adaptors while maintaining the desired semantics.  Such manual work is
not only time-consuming, but exceedingly error-prone.  Moreover,
errors in environment-facing interfaces can not only lead to
corruption of important data, but also to significant security
vulnerabilities.  In order to design and implement survivable,
long-lived, complex software systems that are robust to changes 
in their operational environment---the vision of
the Defense Advanced Research Project Agency's (DARPA) Building
Resource Adaptive Software Systems (BRASS) program---it is necessary to develop new, easier-to-use
and more robust programming systems for managing these ad hoc data
sources as they evolve.

@@

The core result of our research is a new algorithm for
the synthesis of \emph{bijective string lenses}.  Bijective string lenses (\emph{i.e.,}
bidirectional transformations) define a limited set of
transformations between strings.  The domain and range of such
transformations are determined by regular expressions ({\emph{i.e.,}
  regular expressions serve as the types of these transformations).
 In addition, as their name suggests, these transformations are
 \emph{bijections}.
 In other words, the information content of source A is preserved (though usually
rearranged) when data is translated to target B, and vice versa when B
is translated back to A.  Such transformations can rearrange fields of
a record or insert new kinds of syntactic separators (\emph{e.g.,} replacing a
comma with a vertical bar, or the name of one Hypertext Markup
Language (HTML) tag with another)
but they cannot implement more general transformations that elide
irrelevant details, such as the amount of whitespace that separates
two tokens.

@@

\newpage
\section{Methods, Assumptions and Procedures}

@@

Lens-based languages are present in a variety of tools and have found mainstream
industrial use.
Boomerang~\cite{boomerang, Matching10} lenses provide
guarantees on transformations between {\em ad hoc} string document formats.
Augeas~\cite{augeas}, a popular tool that reads Linux system configuration
files, uses the \emph{get} part of a lens to transform configuration
files into a canonical tree representation that users can edit
either manually or 
programmatically.  It uses the lens's \emph{put} to merge the edited
results back into the original string format.  Other lens-based languages and
tools include 
%
GRoundTram~\cite{Hidaka2011GRoundTramAI},
%
BiFluX~\cite{DBLP:conf/ppdp/PachecoZH14}, 
%
BiYacc~\cite{DBLP:conf/staf/ZhuK0SH15},
%
Brul~\cite{DBLP:conf/etaps/ZanLKH16},
%
bidirectional variants of 
relational algebra~\cite{BohannonPierceVaughan},
%
BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
%
spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
graph query languages~\cite{DBLP:conf/icfp/HidakaHIKMN10},
and
Extensible Markup Language (XML) transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.

If $\BRegex$ and $\BRegexAlt$ are regular expressions then
\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
$\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
stochastic regular expressions, so we mark plain REs with overbars.)
We illustrate some of these
combinators by defining lenses on subcomponents of the employee data formats.

@@

% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
%
%\ksf{I don't think this example is necessary.}
%Data can also appear in multiple forms. For example,
%a list of employees with insurance information can either be empty or
%be a sequence of employee/insurance records: \ksf{This example is
%vacuous as you could eliminate the or in favor of the *.  Do somethign
%else to introduce or?}
%
%\begin{lstlisting}
%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
%\end{lstlisting}
The \IterateLens lens is useful for synchronizing a series of items or rows in a
table. For example given a
lens \lstinline{employee_lens} that synchronizes data for a single employee, the
lens
%
\begin{lstlisting}
iterate(id("\n") . employee_lens)  :  ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
\end{lstlisting}
%
transforms a list of employees in the Management format to a list
of employees in the HR format and vice versa. These combinators
are combined in Figure~\ref{fig:example_lens} to construct a complete lens
between the employee formats.

\begin{figure}
\begin{lstlisting}
let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
let employee_lens = name_lens . disconnect(salary,"","unk","") 
                                                    . disconnect("",company,"","UNK") in
let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
ins(header) . employees_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance
\end{lstlisting}
%\ksf{I don't think this is necesssary.}
%let emp_lens = or(id(""),employees_lens) in
  \caption{A lens that synchronizes management and HR employee files}
  \label{fig:example_lens}
  \FS
\end{figure}

@@

  \end{tikzpicture}
  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
    \RegexAlt{}, and examples, \Examples{}, are given as input.
    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
    \RegexAlt{}, and \Examples{}.
    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
    that is equivalent to \DNFLens{}.}
  \label{fig:schematic-diagram-synthesis}
  \FS
\end{figure}

@@

\newpage
\section{Results and Discussion}

@@

\paragraph*{Benchmark Suite Construction}
We constructed our benchmarks primarily by adapting examples from
Augeas~\cite{augeas} and 
Flash Fill~\cite{gulwani-popl-2014}.
%
Augeas is a configuration editing system for Linux that uses lens
combinators similar to those in Boomerang.  
% , with some small
% differences. However, it transforms 
% strings on the left to structured trees on the right rather than
% transforming strings to strings.
% We adapted these Augeas lenses to our setting by converting the
% right-hand sides to strings that correspond to serialized versions
% of the tree formats
We derived 29 of the benchmark tests by
adapting the first 27 lenses from the Augeas library in alphabetical order,
as well as the lenses 
\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
by the `A' lenses.
Furthermore, the 12 last synthesis problems derived
from Augeas were tested after our synthesizer was
finalized, demonstrating that the optimizations were not
over-tuned to perform well on the testing data.
%
Flash Fill is a system that allows users to specify common string
transformations by example~\cite{gulwani-popl-2014}.  
We derived three benchmarks from the first few examples in this
paper and one from the running example on
extracting phone numbers.

@@

Figure~\ref{fig:definition-sizes} shows the complexity of our type-based
specifications as well as our example counts.
An average benchmark has a type-based specification that can be represented using
310 Abstract Syntax Tree (AST) nodes, and requires 1.1 input/output examples.
Our benchmarks vary from simple problems, like changing the
representation of dates
(with a specification size of 85, and a generated lens size of 79), to 
complex tasks, like transforming configuration files for server monitoring
software into dictionary form (with a specification size of 670 and
a generated lens size of 651).  On average, the size of the generated lens is
89\% the size of its type specifications.

@@

We analyzed the same 39 lens synthesis tasks as we did for the original
Optician system.
We also experimented using our tool to synthesize quotient lenses
between XML, Resource Description Framework (RDF) and
Javascript Object Notation (JSON) formats using data from the data.gov
database; the data consisted of census statistics, demographic
statistics, wage comparison data, and crime index data).
Recall that, to use bijective synthesis, we had to modify some of the 
benchmarks (10 of 39) to make them bijective---for instance, by normalizing whitespace.
Because quotient lenses are more expressive, we were able avoid such
modifications. This
experience alone points to the benefits that quotient lens synthesis
brings to the table.

@@

We are interested in comparing three different approaches, which vary
in the amount of synthesis used. 
In the first approach, which we call \QRESize{}, the programmer uses
quotient lens synthesis.  She must write QRE specifications of the source and
target formats and she may give examples.
In the second approach, ``Bijective Synthesis'' (\canonizeAndSpecSize{}), the
programmer uses the Optician bijective synthesis algorithm.
She must write canonizers by hand, along with 
regular expressions to describe the external
representations of the source and target formats. (The internal
formats can be inferred from the canonizers.) She may also
provide examples to help in the synthesis of the bijective lens.
In the third approach, ``No Synthesis'' (\LensAndSpecSize{}), the
programmer writes the lens between the source and target formats
entirely by hand, including the descriptions of the source and target
formats.

For each problem in the benchmark suite, we calculate the following measures
as proxies for the level of programmer effort when using each of the three
approaches:

%
\begin{itemize}
  \item[\QRESize{}:] 
  The number of AST nodes in the QRE specifications for the source and
  target formats, including examples. 
  \item[\canonizeAndSpecSize{}:] 
  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$
  requiring a  non-trivial canonizer (``canonizers'' as transformations that
  project a string to its equivalence class representative; they are written as
  programs in Boomerang), and (3) the number of AST nodes in the
  examples.  We use (1) to estimate the cost of describing
  the external source and target formats and (2) to estimate the
  cost of writing the requisite canonizers
  by hand.  We count the nodes in the examples because they would be
  fed to the bijective synthesizer.  
  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
  automatically generated from the corresponding QRE $q$, and it is
  possible that a human-written version might be smaller.
  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
  number of AST nodes in the synthesized QRE lens.  We use (1) to
  estimate the burdern of describing the source and target formats
  and (2) to estimate the burdern of writing the appropriate lens by
  hand. These counts are also approximations, as
  $W(q)$ and the synthesized lens may be larger than one written by hand.
\end{itemize}

@@

\subsection{Simple Symmetric Synthesis Algorithms}
In addition to the results above, we implemented simple symmetric lenses as an extension to
Boomerang~\cite{boomerang}.  
This allows users
to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into manually-written
lenses, and reference previously defined lenses during synthesis. All experiments
were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of 1600 MHz DDR3
running macOS Mojave.

@@

\begin{figure}
  \centering
  \includegraphics{times}
  \vspace{-2ex}
  \caption{Number of
    benchmarks that can be solved by a given algorithm in a 
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. Without compositional
    synthesis it is able to complete 26. Each benchmark specification includes
    source and target (potentially annotated) regular expressions, and between
    one and three sufficient examples.}
  \label{fig:symm-times}
\end{figure}

For each benchmark in the suite and each mode, we ran the system with a timeout
of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:symm-times}
summarizes the results of these tests. We found that our algorithm is able to
synthesize all of the benchmarks in under 30 seconds. Without compositional
synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.

@@

\newpage
\section{Conclusions}

@@

\newpage

\section{References}

%\printbibliography[heading=none]
\bibliographystyle{apalike}
%\bibliographystyle{IEEEtran}
\bibliography{local.bib,bcp.bib}
